---
title: "FP - IV: Ashraf and Galor (2011) [100 Points]"
subtitle: "ECON 474 Econometrics of Policy Evaluation"
author: "Anahi Rodriguez"
output: 
  html_document:
    toc: yes
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# **[30 Points]** Research Design
********************************************************************************

This section is centered around understanding the research design for this paper.

## **[10 Points]** Research Question
************************************************************

What is the research question of this paper?

*Hint: economists are notorious for not explicitly stating their research question in their papers.*

**Your answer:**
Asks if the Malthusian theory is true; Do improvements in technology and higher land productivity affect population density or the standard of living



## **[20 Points]** Empirical Design
************************************************************

1. What is the casual technique of this paper, and why is it chosen?

**Your answer:**
Instrumental Variables and it is chosen becuase there is exogenous sources of variation across the country in technological advancements and land prductivity


2. Name one possible threat to identification.

**Your answer:**
It could be that the technological advancements of one area spillover into those sorroudning areas, violating the SUTVA assumption


# **[70 Points]**  Replication
********************************************************************************

The paper estimates the regression:

$$ln(P_{i,t}) = \alpha_{0,t} + \alpha_{1,t} \widehat{ln(T_i)} + \alpha_{2,t} ln(X_i) + \alpha_{3,t} \Gamma_i + \alpha_{4,t} D_i + \epsilon_{i,t}$$
$$ln(T_i) = \rho_{0,t} + \rho_{1,t} Z_i + \rho_{2,t} ln(X_i) + \rho_{3,t} \Gamma_i + \rho_{4,t} D_i + \xi_{i,t}$$

separately for different years $t$, where $i$ is the modern day country.

The mapping of the variables in the regression to the data are:

- $P_{i,1500}$ - `ln_pd1500` (population density in 1500 CE)
- $T_i$ - `ln_yst` (years since neolithic period)
- $X_i$ - `pc_lnar_lnas` (land productivity)
- $\Gamma_i$ - geographical controls for country $i$
    - `ln_abs_lat` - absolute latitude
    - `distcr1000` - mean distance to nearest coast or river
    - `land100cr` - percentage of land within 100 km of coast or river
- $D_i$ - `continent`
- $Z_i$ - instruments
    - `animals` - number of domesticable animals species
    - `plants` - number of domesticable plants species


**[5 Points]** Load in the data and any necessary packages:

```{r loading please wait}
library(tidyverse)
library(fixest)
library(haven)

setwd("~/Desktop/data econ 474")
df = read_dta("20081371_Dataset.dta")


```


## **[30 Points]** Table 2: Population Density in 1500 CE
************************************************************

Here you are going to somewhat replicate a modified table 2.
Specifically, you are going to replicate columns 5 and 6 and add the reduced form estimates.
There is a typo in the paper's code, so your estimates will be slightly different but the number of observations will be the same.
For some guidance, the OLS estimate on `ln_yst` should be 1.375, the RF estimate on `animals` should be 0.0659, and the 2SLS estimate on `ln_yst` should be 2.051.

For each regression, use the subset of the data where `cleanhibbs == 1` and use heteroskedastic robust standard errors.

Put all three regressions into an `etable()`.
Rename all variables appropriately (recall you cannot rename the cluster variable).
Specify the `fitstat` argument to match.

```{r tab2}
#`feols(outcome ~ exogenous variables | endogenous variables ~ instrumental variables)`



table_2_col5 = feols(ln_pd1500 ~ ln_yst + pc_lnar_lnas + ln_abslat + distcr1000 + land100cr | continent,
                data = subset(df, cleanhibbs ==1),
                se = 'hetero')


table_2_col6 = feols(ln_pd1500 ~  pc_lnar_lnas + ln_abslat + distcr1000 +
                 land100cr  | continent | ln_yst ~ animals + plants,
                data = subset(df, cleanhibbs ==1),
                se = 'hetero')


reduced_form_fit = feols(ln_pd1500 ~ animals + plants + ln_yst + pc_lnar_lnas + ln_abslat + distcr1000 + land100cr | continent,
                         data = subset(df, cleanhibbs ==1),
                         se = 'hetero')


tab_dict = c(ln_yst = "Log years since Neolithic transition",
             pc_lnar_lnas = 'Log land productivity',
             ln_abslat = 'Log absolute latitude',
             distcr1000 = 'Mean distance to nearest coast or river',
             land100cr = 'Percentage of land within 100 km of coast or river',
             'animals' = 'Animals',
             'plants' = 'Plants')


etable(table_2_col5, table_2_col6, reduced_form_fit,
       fitstat = c('r2', 'n'),
       se.below = T,
       dict = tab_dict)




```


## **[35 Points]** Frisch-Waugh-Lovell - Figure 3b
************************************************************


Use the Frisch-Waugh-Lovell theorem to plot the partial relationship between `ln_pd1500` and `pc_lnar_lnas` with `ggplot()` using the OLS regression from above, but with the subset of the data `cleanpd1500 == 1`.
Set the scatter plot point type to represent the continents, and use `geom_text()` to label all of the points with country `code`s.
Label the axes and title appropriately. 

*Hint: get the variables* `code` *and* `continent` *from the matching subset of the original data set.*


```{r fig3b}
ggplot(data = subset(df, cleanpd1500 ==1),
       aes(x = pc_lnar_lnas, y = ln_pd1500 )) +
  geom_point(aes(shape = continent)) +
  scale_shape_manual(values = c(16, 1, 22, 18, 24)) +
  geom_smooth(method = 'lm', se = F, color = 'black') +
  geom_text(aes(label = code ), size = 2, vjust = 2) + 
  labs(title = 'Panel B. The Partial Effect of Land Productivity on Population 
       Density in 1500 CE', x = 'Log years since transition',
       y = 'Log Population density in 1500 CE') +
  theme_bw() +
  scale_x_continuous(breaks = NULL) + 
theme(legend.position = c(.325,.92),
        legend.direction = 'horizontal',
      legend.title = element_blank())

```